---
title: "Unit 2: Statistical analysis for ZWE dataset"
author: "Esteban Correa"
date: "November, 2019"
output: rmarkdown::github_document

---

# Summary 

Our analysis pipeline can be summarized as follows:

* Create the survey design object 
* Perform population characteristics for table 1
* Perform unadjusted analysis 
* Perform adjusted analysis 

# Load datasets

Let's load our prepared dataset

```{r setup, include=TRUE, echo=TRUE, warning=FALSE,comment=FALSE,message=FALSE}
rm(list = ls())
library(foreign)
library(plyr)
library(dplyr)
library(survey)
library(jtools)
library(kableExtra)
library(broom)
library(ggstance)
# load female, male and hiv records of Zimbabwe
load("../data/zwe_dataset.RData")
summary(allZWE)
```

## Create survey design

We create the survey object using survey package from Lumley et al. It requires the ID to identify the PSU, the strata and the individual weights. 

```{r}
# We need to set this flag, because we do not have variance within psu
options (survey.lonely.psu = "remove")
dhs1<- svydesign(id = ~PSU, 
                 strata = ~V022,
                 weights = ~indivweight,
                 nest=TRUE,
                 data = allZWE)

```

## Descriptive statistics

This is one way to create your table 1 manually, however there is another using table1 package. The idea is to create contigency tables per each variable

```{r}
mymr<-2
round(svytable(~migstatus, dhs1),1)
svyttest(migstatus~0, dhs1)

round(prop.table(svytable(~hivStatus+migstatus, dhs1),margin = mymr)*100,0)
summary(svytable(~hivStatus+migstatus, dhs1), statistic="Chisq")


round(prop.table(svytable(~sex+migstatus, dhs1),margin = mymr)*100,0)
summary(svytable(~sex+migstatus, dhs1), statistic="Chisq")

svyby(~age, ~migstatus, dhs1, svymean, se=T,ci=TRUE,vartype =c("se","ci","cv","cvpct","var"))
svyboxplot(age~migstatus,dhs1)
svyttest(age~migstatus,dhs1)

round(prop.table(svytable(~agegroup+migstatus, dhs1),margin = mymr)*100,0)
summary(svytable(~agegroup+migstatus, dhs1), statistic="Chisq")

round(prop.table(svytable(~education+migstatus, dhs1),margin = mymr)*100,0)
summary(svytable(~education+migstatus, dhs1), statistic="Chisq")

round(prop.table(svytable(~Wealthindexcombined+migstatus, dhs1),margin = mymr)*100,0)
summary(svytable(~Wealthindexcombined+migstatus, dhs1), statistic="Chisq")

round(prop.table(svytable(~healthInsured+migstatus, dhs1),margin = mymr)*100,0)
summary(svytable(~healthInsured+migstatus, dhs1), statistic="Chisq")

round(prop.table(svytable(~maritalStatus+migstatus, dhs1),margin = mymr)*100,0)
summary(svytable(~maritalStatus+migstatus, dhs1), statistic="Chisq")

# round(svytable(~religion+migstatus, dhs1),1)
# round(prop.table(svytable(~religion+migstatus, dhs1),margin = mymr),0)*100
# summary(svytable(~religion+migstatus, dhs1), statistic="Chisq")

round(prop.table(svytable(~placeType+migstatus, dhs1),margin = mymr)*100,0)
summary(svytable(~placeType+migstatus, dhs1), statistic="Chisq")

round(prop.table(svytable(~condomLastTime+migstatus, dhs1),margin = mymr)*100,0)
summary(svytable(~condomLastTime+migstatus, dhs1), statistic="Chisq")

round(prop.table(svytable(~previousSTI+migstatus, dhs1),margin = mymr)*100,0)
summary(svytable(~previousSTI+migstatus, dhs1), statistic="Chisq")

round(prop.table(svytable(~hivtested+migstatus, dhs1),margin = mymr)*100,0)
summary(svytable(~hivtested+migstatus, dhs1), statistic="Chisq")

round(prop.table(svytable(~sexPartnersLast12+migstatus, dhs1),margin = mymr)*100,0)
summary(svytable(~sexPartnersLast12+migstatus, dhs1), statistic="Chisq")

round(prop.table(svytable(~lifetimePartners+migstatus, dhs1),margin = mymr)*100,0)
summary(svytable(~lifetimePartners+migstatus, dhs1), statistic="Chisq")

round(prop.table(svytable(~Distance2Borders+migstatus, dhs1),margin = mymr)*100,0)
summary(svytable(~Distance2Borders+migstatus, dhs1), statistic="Chisq")

round(prop.table(svytable(~TravelTimes+migstatus, dhs1),margin = mymr)*100,0)
summary(svytable(~TravelTimes+migstatus, dhs1), statistic="Chisq")
```

## Unadjusted analysis

```{r, eval=TRUE}
fitall<-svyglm(migstatus~hivStatus,
            data=allZWE,
            family=quasibinomial(),
            design=dhs1)
summ(fitall, exp = T,confint = TRUE,digits = 2)        

fitall<-svyglm(migstatus~sex,
            data=allZWE,
            family=quasibinomial(),
            design=dhs1)
summ(fitall, exp = T,confint = TRUE,digits = 2)  

fitall<-svyglm(migstatus~
              agegroup,
            data=allZWE,
            family=quasibinomial(),
            design=dhs1)
summ(fitall, exp = T,confint = TRUE,digits = 2)

fitall<-svyglm(migstatus~
              education,
            data=allZWE,
            family=quasibinomial(),
            design=dhs1)
summ(fitall, exp = T,confint = TRUE,digits = 2)

fitall<-svyglm(migstatus~Wealthindexcombined,
            data=allZWE,
            family=quasibinomial(),
            design=dhs1)
summ(fitall, exp = T,confint = TRUE,digits = 2)

fitall<-svyglm(migstatus~healthInsured,
            data=allZWE,
            family=quasibinomial(),
            design=dhs1)
summ(fitall, exp = T,confint = TRUE,digits = 2)

fitall<-svyglm(migstatus~maritalStatus,
            data=allZWE,
            family=quasibinomial(),
            design=dhs1)
summ(fitall, exp = T,confint = TRUE,digits = 2)

fitall<-svyglm(migstatus~placeType,
            data=allZWE,
            family=quasibinomial(),
            design=dhs1)
summ(fitall, exp = T,confint = TRUE,digits = 2)
               
fitall<-svyglm(migstatus~condomLastTime,
            data=allZWE,
            family=quasibinomial(),
            design=dhs1)
summ(fitall, exp = T,confint = TRUE,digits = 2)

fitall<-svyglm(migstatus~
              previousSTI,
            data=allZWE,
            family=quasibinomial(),
            design=dhs1)
summ(fitall, exp = T,confint = TRUE,digits = 2)

fitall<-svyglm(migstatus~
              hivtested,
            data=allZWE,
            family=quasibinomial(),
            design=dhs1)
summ(fitall, exp = T,confint = TRUE,digits = 2)

fitall<-svyglm(migstatus~
              sexPartnersLast12,
            data=allZWE,
            family=quasibinomial(),
            design=dhs1)
summ(fitall, exp = T,confint = TRUE,digits = 2)

fitall<-svyglm(migstatus~
              lifetimePartners,
            data=allZWE,
            family=quasibinomial(),
            design=dhs1)
summ(fitall, exp = T,confint = TRUE,digits = 2)

fitall<-svyglm(migstatus~
              Distance2Borders,
            data=allZWE,
            family=quasibinomial(),
            design=dhs1)
summ(fitall, exp = T,confint = TRUE,digits = 2)

fitall<-svyglm(migstatus~
              TravelTimes,
            data=allZWE,
            family=quasibinomial(),
            design=dhs1)
summ(fitall, exp = T,confint = TRUE,digits = 2)

```

## Multivariable analysis

We make use of jtools to extract our beta coefficients in exponential form and with the VIF for multicollinearity checking.

```{r, results="asis"}
fitall<-svyglm(migstatus~hivStatus+
              sex+
              agegroup+
              education+
              Wealthindexcombined+
              healthInsured+
              maritalStatus+
              placeType+
              condomLastTime+
              previousSTI+
              hivtested+
              sexPartnersLast12+
              lifetimePartners+
              Distance2Borders+
              TravelTimes,
            data=allZWE,
            family=quasibinomial(),
            design=dhs1)
summ(fitall, exp = T,confint = T,digits = 2,model.info = T,model.fit = T,vifs = T)

```

## References

- Thomas Lumley. Complex Surveys: A Guide to Analysis Using R. ISBN: 978-0-470-28430-8. 
